---
title: "Dplyr Manipulate Variables"
author: "Victoire Migashane"
date: "9/5/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(dplyr)
library(palmerpenguins)
df <- penguins
```

#### Dataset

```{r}
head(df)
```
### 1. Pull Function

The `pull()` function lates you extract column values as a vector by name or index. `pull()` is very similar to `$`

#### Arguments

    .data: A data frame to full from.
    var: a variable (column name, positive/negative integer)
    name: Specifies the column to be used as names for a named vector.
    ...: for use by methods.
  
#### Usage

```{r}
df %>% 
  pull(species) %>% # not a data frame but the column cell values
  unique()

#similarly the $
df$species %>% 
  unique()
```
### 2. Relocate Function

The `relocate()` function is use to move a specified column to a different position in the data frame.

#### Arguments

    .data: A data frame to select from.
    ...: columns to move.
    .before: destination of the columns.
    .after: selected using the columns to move
  
#### Usage

```{r}
df %>% 
  relocate(year, sex) %>% 
  head(3)
```

### 3. Select Function

The `select()` function makes it easier to refer (select) variables by name or heper function.

Ex: a:f == Select all columns from `a` on the left to `f` on the right.

#### Arguments:

    .data: A data frame to select from
    ...: One or more unquoted expressions separated by commas.
    
#### Helper function:

    `:`: selecting a range of consecutive variables (select between)
    `-` : Select columns except
    `!`: take the complement of a set of variables
    `c()`: for combining selections.
    `& and |`: for selecting the intersextion of twe sets of variables
    `everything()` : Select every column.
    `last_col()`: Select the last column.
    `starts_with()`: Select columns whose names starts with (characture string).
    `ends_with()`: Select columns whose name ends with (characture string).
    `contains()`: Select columns whose name contains (characture string).
    `matches()`: Select columns whose name matches a regular expression.
    `num_range()`: Select columns that mache a numerical range like x1,x2,x3
    `all_of(), any_of()`: Select columns whose names are all present in a character vector.
    `one_of()`: select columns whose names are in a group of names.
    `where()`: select a variable with a function

#### Usage:

```{r}
# Return every column between bill_length_mm and body_mass_g
df %>%
    select(bill_length_mm:body_mass_g) %>%
    head(3)

# return every column that is not in the list
df %>% 
    select(-c(island, species, sex, year)) %>%
    head(3)
# return every column
df %>%
    select(everything()) %>%
    head(3)

# return the last column
df %>% 
    select(last_col()) %>%
    head(3)
```
```{r}
# return every column that starts with b and ends with m
df %>%
    select(starts_with("b") & ends_with("m")) %>%
    head(3)

# return column names that contain the string land
df %>%
    select(contains("land")) %>% 
    head(3)

# return columns whose names are all present in a character vector
# also works for any_of()
df %>% 
    select(all_of(c("bill_length_mm", "flipper_length_mm", "body_mass_g"))) %>% 
    head(3)

# return columns whoes data type are integers
df %>% 
    select(where(is.numeric)) %>% 
    head(3)
```

