---
title: "Dplyr Make New Variables (Columns)"
author: "Victoire Migashane"
date: "9/6/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(palmerpenguins)
df <- penguins
```
```{r}
head(df)
```

### 1. Rename function

The `rename()` function changes the names of individual columns using the `new_name = old_name`. A similar function is `rename_with` which renames the column using a function.

#### Arguments

    .data: A data frame to apply changes to
    ...: rename selected volumns `new = old`
    .fn: transforms the selected columns to return a charcture vector of same length
    .cols: columns to rename (deafult is ALL the columns)
    
#### Usage

```{r}
df %>% 
    rename(gender = sex, col = c("bill_length_mm", "bill_depth_mm", "flipper_length_mm")) %>% 
    head(3)
```


### 2. Mutate and Transmute Functions

The `mutate()` function is used to acompute and append new columns to an existing data frame.

The `transmute()` function does the opposite of `mutate()`. This function add new variables and drops (remove) existing ones.

#### Arguments

    .data: A data frame
    ...: The name given to the new column
    .keep: control which columns to retain in the out put. `.keep = c("all", "used", "unused", "none")`
    .before/.after: control where new columns should appear

#### Usage

```{r}
df %>% 
    mutate(bill_and_flipper_len = bill_length_mm + flipper_length_mm, .keep = "used") %>% 
    head(3)

df %>% 
    transmute(species, island, bill_and_flipper_len = bill_length_mm + flipper_length_mm) %>% 
    head(3)
```


### 3. Mutate window functions

The `muate()` function uses uses window functions. These are functions that take vector of values and return another vector of values with same lenght.

#### Offset

    `lag()`: offset elements by 1
    `lead(): offset elements by -1
    
#### Cumulative aggregate

    `cumall()`: cumulative all
    `cumany()`: cumulative any
    `cummax()`: cumulative max
    `cummin()`: cumulative min
    `cummean()`: cumulative mean
    `cumprod()`: cumulative prod
    `cumsum()`: cumulative sum

#### Ranking

    `cume_dist()`: proportion of all values <=
    `dense_rank()`: randk w ties  = min, no gapes
    `min_rank()`: randk with ties = min
    `ntile()`: bins into n bins
    `percent_rank()`: min_rank scaled to [0,1]
    `row_number()`: rank with ties =  "first
    
#### Math

    arithmetic ops: `+, -, *, /, ^, %/%, %%`
    logs: `log(), log2(), log10()`
    logical comparizons: `<, >, <=, >=, ==, !=`
    x >= left <= right: `between()`
    sage == for floating point numbers: `near()`
    
#### Miscellaneous

    `case_when()`: multi_case if_else()
    `coalsce()`: first non-NA values by element across a set of vector
    `if_else()`: element-wise if() + else()
    `na_if`: replace specific values with NA
    `pmax()`: element wise max()
    `pmin()`: element wise min()


