---
title: "Dplyr Subset Observations"
author: "Migashane Victoire"
output: html_notebook
---

```{r}
library(palmerpenguins)

data(penguins)
df <- penguins
head(df)
```

### 1. Filter Function

The `filter()` function is used to extract rows of a data frame that satisfy logical conditions.

Arguments:

    .data: A data frame to extract from
    ...: Expressions that return a logical value, adn defined in terms of the variables in the data frame.
    .preserve: if FALSE (deafult) the grouping structure is recalculated based on the resulting data.

Useful functions:

    == : equal to, != : not equal to,
    > : greater than, < : less than,
    >= : greater than or equal to, 
    <= : less than or eual to,
    & : and, | : or, ! : not, 
    %in% = in, xor(): or, 
    is.na() : is none, !is.na() isn't none, 
    between(): is between, near(): is near
    
Usage:

```{r}
df %>%
    filter(species == "Adelie" & sex == "male" & between(bill_length_mm, 38,39)) 

df %>%
    filter(between(flipper_length_mm, 180,190) | is.na(bill_depth_mm),sort = TRUE) %>%
    count(species)
```

### 2. Slice Function

The `slice()` function lets you select rows by their index positions.

Arguments:
    
    .data: A data frame to select from.
    ...: the index positions (integer row values)
    .preserve(): False: false the grouping structure is recalculated based on the resulting data, other wise the groupping is kept the same.
    order_by: the variable(column) to order by.
    with_ties: if TRUE return more rows than you request, other wise return one row.
    weight_by: must evaluate a vector of positive numbers an that all sum to 100%.
    repace: if TRUE, sampling should be performed with replacement.
    
Related slice methodes:

    slice_head(): Returns the first few rows.
    slice_tail(): Returns the last few rows.
    slice_min(): Returns the minimum value in the specified column
    slice_max(): Returns the maximum value in teh specified column
    slice_sample(): Return the sample of specified lengh

Usage:

```{r}
# return row 5 to row 10
df %>% slice(5:10)

# return the first 3 rows
df %>% slice_head(n = 3)

# return the random 8 rows
df %>% slice_sample(n = 8)

# return the min value in a column
df %>% slice_min(body_mass_g )
```

### 3. Distict Function

The `distinct()` function selects only unique rows from from a data frame, removing all rows with duplicated values. This is similar to unique() but faster.

Arguments:

    .data: A data frame to select from.
    ...: Optional variable to use when determining uniqueness.
    .keep_all: if TRUE, keep all variables in the data frame if ... is not distinct
Usage:

```{r}
df %>%
    distinct(island)

# similar to distinct is count()
df %>%
    count(island)
```


### 4. Sample Function

The  `sample()` function retunrn a random sample of specified zise from the elements of argument x. `sample()` takes the place for `sample_frac()` and `sample_n()`. 

Arguments:

    x: A vector of one or more elements or a positive integer value.
    n: The number of items to choose form.
    size: The number of items to return.
    replace: if TRUE, sampling should be performed with replacement.
    prob: A vector of probability weights to get the elements of the sample vector.
    useHash: indicats if the hash-version of the algorithm should be used.
    
Note: Install `sampling` package from CRAN for other methods of weighted sampling.

Usage:

```{r}
sample(c(0,1),10,  size = 4)
```
