---
title: "Dplyr Data Summary"
outhor: "Migashane Victoire"
output: html_notebook
---

### Install and load dplyr
```{r}
# To get more data sets, you can use `library(help = "datasets")`

rm(list = ls())
# install.packages(dplyr)
library(dplyr)

# 1st way to load R data
#force iris to a tibble (lazy data frame, doesn't do all data frame operations)
data <- dplyr::tbl_df(iris)
head(data,3)

#2nd way to load R data
# `data()` loads specific data sets (good data frame)
data(iris)
df <- iris
head(df)

# 3rd way to load data
# Read a file in table formatand create a data frame from it.
# my_data <- read.csv("my_data.csv")
```

### View data
```{r}
# View data set in spreadsheet-like format (note: capital V)
View(df)

# Shows the structure of an arbitrary R object
str(df)

# For information dansity use `glimpse()`, this will show as much data as possible similar to str()
dplyr::glimpse(df)

# `%>%` Passing the object to the left as the first argument of the right function
df %>% summary()
```

### Summarise Data

#### 1. Summarise or Summarize Function
Summary functions take vectors as input and return a single value. Apply summary functions to columns to create a new table of summary statistics.

Arguments:
    
    .data: A data frame.
    ...: Name value pair of summary functions.
    .groups: grouping structure of the result.
    
Useful functions:

    Center: mean(), median()
    Spread: sd(), IQR(), mad()
    Range: min(), max(), quantile()
    Position: first(), last(), nth(),
    Count: n(), n_distinct()
    Logical: any(), all()
    
Helper functions:
    
    summarise_each(): applies summary function to each column
    count()
    
List of summary functions:

Note: The all take a vector of values and return a single value... useful as summaise() arguments.

    Count:
    
        n(): return number of values/rows
        n_distinct(): return number of unique values
        sum(!is.na()): return number non-NA values
        
    Position:
    
        mean(): return mrean,  also mean(!is.na())
        median(): return the median
        
    Logical:
    
        mean(): proportion of TRUE's
        sum(): number of TRUE's
        
    Order:
    
        first(): first value
        last(): last value
        nt(): value in th nth location of vector
        
    Rank:
    
        quantile(): return nth quantile
        min(): return minimum value
        max(): return maximum value
        
    Spread:
    
        IQR(): inter-quartile range
        mad(): median absolute deviation
        sd(): standard deviation
        var(): variance
    
Usage:
```{r}
# summarise()
df %>% 
    summarise(ave = mean(Sepal.Length), len = n()) # n() return the dataset size

# simular function (summarize())
df %>% 
    summarize(firstLen = first(Sepal.Length), Is_NA = sum(is.na(Sepal.Length)), SD = sd(Sepal.Length)) #sd() return the standard deviation

# summarise_each()
df %>% summarise_each(funs(mean(Sepal.Width)))
```

#### 2. Count Function

count() function returns the count of unique value for one or more variables.

df %>% count(a, b) is roughly equivalent to df %>% group_by(a, b) %>% summarise(n = n()).

Arguments:
    
    .data
    wt = NULL: Frequency weights. Can be NULL or a variable.
    sort = FALSE: If TRUE, will show the largest groups at the top.
    name = "ne_column": The name of the new column in the output. (column n)
    .drop = FALSE: Ff FALSE will include counts for empty groups
    
Usage:
```{r}
df %>% count(Species, name = "Unique_count")
```
