---
title: "Dplyr Summary and Group Functions"
output: html_notebook
---

### Install Packages
```{r}
#Install the dplyr package
#install.packages("dplyr")

# Load the dplyr library into my current session
library(dplyr)

# This package contains a variety of datasets. For a complete list
# library(help = "datasets")
# library(datasets)  # practic R datasets

library(palmerpenguins) # onther datase
```

### Load R datasets
```{r}
# load the data
data(iris)
data(penguins)

df <- iris
data <- penguins

# To load from csv 
# df <- read.csv("yourData.csv")
```

### iris dataset and penguins dataset
```{r}
head(df,3)
```
```{r}
head(data,3)
```

### Summary cases
Summary functions take vectors as input and return one value. Apply summary functions to columns to create a new table of summary statistics.

### 1. summarise, summarize and summary

Some useful arguments for `r "summarize()"`:
    
    .data
    Center: mean(), median()
    Spread: sd(), IQR(), mad()
    Range: min(), max(), quantile()
    Position: first(), last(), nth(),
    Count: n(), n_distinct()
    Logical: any(), all()
    
How to use summary functions:
```{r}
# summarise()
df %>% 
    summarise(mean(Sepal.Length), n()) # n() return the dataset size
```
```{r}
# simular function (summarize())
df %>% 
    summarize(first(Sepal.Length), any(), sd(Sepal.Length)) #sd() return the standard deviation
```
```{r}
# get a general summary
df %>% 
    summary()
```

#### 2. Count Function
(
`r "count()"` function returns the count of unique value for one or more variables.

Some useful arguments for `r "count()"`:

    .data
    wt = NULL: Frequency weights. Can be NULL or a variable.
    sort = FALSE: If TRUE, will show the largest groups at the top.
    name = "ne_column": The name of the new column in the output. (column n)
    .drop = FALSE: Ff FALSE will include counts for empty groups
    
How to use `r "count()"`
```{r}
data %>% count(species, sort = TRUE, name = "count")
```

### Group Cases
### 1. Group By Function

`r "group_by()"` function is used to create a grouped copy of a table by columns. `r "group_by()"` takes an existing tbl and converts it into a grouped tbl where 
operations are performed "by group"

some useful arguments for `r "group_by()"`:

    .data
    .add = FALSE:
    .drop = TRUE:
    
How to use `r "group_by()"`.
```{r}
data %>% 
    group_by(species, sex) %>% count(species, name= "species count")
```

### 2. Row Wise Function
`r "rowwise()"` function is used to group data into individual rows. `r "rowwise()"` works like `r "group_by()"` in the sense that it doesn’t change what the data looks like; it changes how dplyr verbs operate on the data.

    * use with `r "mutate()"` to maximaze results.
    
How to use `r `rowwise()`.
```{r}
data %>% 
    rowwise(species) %>% 
    count(species,name = "count of species")
```